An Improved Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium

### 2020-12-31 simplified_sdc_intro/paper.tex
more details

With SDC evolution, when we are in NSE, we need to do the advective
and reactive updates together.  With the tabulated NSE reaction
values, we get instantaneous values of the reaction sources, but the only
time derivative we have is for $Y_e$.  We will
use a basic predictor-corrector update:
\begin{equation}
\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \left [\Rb (\Uc) \right ]^{n+1/2}
\end{equation}
Calling the NSE table on the starting state gives us:
\begin{equation}
\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
\end{equation}
To get the reactive source, we need to use the iteratively-lagged time-centered information
from $\Ic_q^{n+1/2,(k-1)}$.  We only need to update the internal energy and auxiliary data,
since together with the density, that is all that is needed for the equation of state to get the
updated temperature.    So our sources are:
\begin{equation}
\Rb(\rho e) \approx \Ics_q(\rho e)
\end{equation}
and
\begin{equation}
\Rb(\rho \alpha_l) \approx \rho^{n+1/2} \Ics_q(\alpha_l)
\end{equation}
were $\rho^{n+1/2} = (\rho^n + \rho^\star)/2$.

Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
$\Rb(\Uc^n)$ as: \MarginPar{it might be better if we had $\Ics_\Uc$}
\begin{equation}
R(\rho Y_e) = \rho^{n+1/2} (dY_e/dt)^n
\end{equation}
which gives an update for the electron fraction as:
\begin{equation}
(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho^{n+1/2} (dY_e/dt)^n
\end{equation}
From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
$e^\star$, which allows us to call the EOS and get an updated
temperature, $T^\star$. \MarginPar{Andy suggests time-centering $dY_e/dt$}
We can then again call the NSE table:
\begin{equation}
\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
\end{equation}
where we use the outputs to define the new values of the composition variables.
Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
as their new values:
\begin{align}
\rho^{n+1,(k)} &= \rho^\star \\
(\rho u)^{n+1,(k)} &= (\rho u)^\star
\end{align}
We then compute the energy release as
\begin{equation}
       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
\end{equation}\MarginPar{is this right?}
and then do the final update of the total energy:
\begin{equation}
(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
\end{equation}

>>>>

With SDC evolution, when we are in NSE, we need to do the advective
and reactive updates together.  With the tabulated NSE reaction
values, we get instantaneous values of the reaction sources, but the only
time derivative we have is for $Y_e$.  We will
use a basic predictor-corrector update:
\begin{equation}
\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \left [\Rb (\Uc) \right ]^{n+1/2}
\end{equation}
Calling the NSE table on the starting state gives us:
\begin{equation}
\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
\end{equation}
To get the reactive source, we need to use the iteratively-lagged time-centered information
from $\Ic_q^{n+1/2,(k-1)}$.  We only need to update the internal energy and auxiliary data,
since together with the density, that is all that is needed for the equation of state to get the
updated temperature.    So our sources are:
\begin{equation}
R(\rho e) \approx \Ics_q(\rho e)
\end{equation}
and
\begin{equation}
R(\rho \alpha_l) \approx \rho^{n+1/2} \Ics_q(\alpha_l)
\end{equation}
where $\rho^{n+1/2} = (\rho^n + \rho^\star)/2$.

Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
$\Rb(\Uc^n)$ as: \MarginPar{it might be better if we had $\Ics_\Uc$}
\begin{equation}
R(\rho Y_e) = \rho^{n+1/2} (dY_e/dt)^n
\end{equation}
which gives an update for the electron fraction as:
\begin{equation}
(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho^{n+1/2} (dY_e/dt)^n
\end{equation}
We can then get $\alpha_l^\star = (\rho \alpha_l)^\star/\rho^\star$.
From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
$e^\star$, which allows us to call the EOS and get an updated
temperature, $T^\star$. \MarginPar{Andy suggests time-centering $dY_e/dt$}
\begin{equation}
T^\star = T(\rho^\star, \alpha_l^\star, e^\star)
\end{equation}
We can then again call the NSE table:
\begin{equation}
\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
\end{equation}
where we use the outputs to define the new values of the composition variables.
Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
as their new values:
\begin{align}
\rho^{n+1,(k)} &= \rho^\star \\
(\rho u)^{n+1,(k)} &= (\rho u)^\star
\end{align}
We then compute the energy release as
\begin{equation}
       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
\end{equation}\MarginPar{is this right?}
and then do the final update of the total energy:
\begin{equation}
(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + \Delta t (\rho \Sdot)^{n+1/2}
\end{equation}
